Supplementary Information

### 2020-08-27 doc/manuscript/MEE_manuscript.Rmd
update SI fig numbers

We analyzed previously collected tree-ring data from ten sites ranging from 9.15$^\circ$ to 61.30$^\circ$ N latitude and representing a wide range of forest and tree types: tropical broadleaf deciduous and evergreen, temperate broadleaf deciduous and needleleaf evergreen, and boreal needleleaf evergreen (Tables 1, S1-S2).
Nine of these sites (exception: LT) are large forest dynamics plots of the Forest Global Earth Observatory [ForestGEO; @anderson-teixeira_ctfs-forestgeo:_2015].
Trees of species represented within the ForestGEO plots were cored within the plot (n=#) and/or in the vicinity (n=#; max distance: # km), following a variety of sampling protocols designed to meet the varied objectives of the original studies (Table S1). 
In using this variety of data sources, we ensured that our approach was able to handle a variety of challenges presented by varying methodologies and forest types.

@@

For each year in the tree-ring records, we reconstructed $DBH$, as detailed in *Appendix S1*. 
In most cases, when a recent $DBH$ measurement was available, $DBH$ was reconstructed from the outside in. 
In cases where $DBH$ was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings (DENDORO_REF), $DBH$ was reconstructed from the inside out. 
In either case, we used allometric equations between $DBH$ and bark thickness to account for changes in bark thickness as the tree grew (Appendix S1; Table S2).

@@

![**Figure 2 | Example comparison of climate sensitivity derived via traditional methods (**a**) and our approach (**b-f**).** Example is for the sensitivity of 14 species at SCBI (codes given in Table S1) to potential evapotranspiration (PET), identified by both traditional methods and our method to be among the top climate drivers. Panel (**a**) shows a matrix of Pearson correlations between ring- width index and monthly climate variables. Panels (**b-d**) give statistics for time windows tested in *climwin*, where window open and close indicate months prior to current August, and cells across the lower diaganol indicate single-month tests (akin to panel **a**). Panels (**b**) and (**c**) give values of linear and quadratic terms for each time window, and (**d**) gives the $\Delta AIC$ for each. The time window with the minimum $\Delta AIC$ (0-3 months prior to August, or May-July; black circles), was identified as the optimal window. Panel (**e**) shows the correlation of individual-level residuals to PET, with the function fit in *climwin*. Finally, panel (**f**) shows GLS model output, where $PET$ was a candidate driver variable (along with $PPT$; $DBH$ not included in this model). Plotted are responses of species for which $PET$ was included in the top model, with best-fit polynomials plotted with solid lines when both first- and second-order terms are signficant, DASHED lines when only one term is signficant, and DOTTED lines when neither is signficant. Transparent ribbons indicate 95% confidence intervals.](tables_figures/quilt_comparison.png)

@@

**Interactive effects of $DBH$ and climate were found for #% of species-variable combinations**. 
Interactions were significant for over half of $DBH$-precipitation variable interactions for all three growth metrics (*51% for all*; Table S4).
The majority (#%) of these interactions were positive, indicating that larger trees generally respond more positively (or less negatively) to precipitation or it's frequency (Fig. **6?**).
Interactions between temperature variables and $DBH$ were significant for *~40%* of cases considered (Table S4). 
The majority (#%) of these interactions were positive, indicating that larger trees generally respond less negatively (or more positively) to $T_{max}$ or $PET$ (Fig. **6?**).

>>>>

We analyzed previously collected tree-ring data from ten sites ranging from 9.15$^\circ$ to 61.30$^\circ$ N latitude and representing a wide range of forest and tree types: tropical broadleaf deciduous and evergreen, temperate broadleaf deciduous and needleleaf evergreen, and boreal needleleaf evergreen (Tables 1, S1-S2).
Nine of these sites (exception: LT) are large forest dynamics plots of the Forest Global Earth Observatory [ForestGEO; @anderson-teixeira_ctfs-forestgeo:_2015].
Trees of species represented within the ForestGEO plots were cored within the plot (n=#) and/or in the vicinity (n=#; max distance: # km), following a variety of sampling protocols designed to meet the varied objectives of the original studies (Tables S1, S3). 
In using this variety of data sources, we ensured that our approach was able to handle a variety of challenges presented by varying methodologies and forest types.

@@

For each year in the tree-ring records, we reconstructed $DBH$, as detailed in *Appendix S1*. 
In most cases, when a recent $DBH$ measurement was available, $DBH$ was reconstructed from the outside in. 
In cases where $DBH$ was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings (DENDORO_REF), $DBH$ was reconstructed from the inside out. 
In either case, we used allometric equations between $DBH$ and bark thickness to account for changes in bark thickness as the tree grew (Appendix S1; Table S4).

@@

![**Figure 2 | Example comparison of climate sensitivity derived via traditional methods (**a**) and our approach (**b-f**).** Example is for the sensitivity of 14 species at SCBI (codes given in Table S2) to potential evapotranspiration (PET), identified by both traditional methods and our method to be among the top climate drivers. Panel (**a**) shows a matrix of Pearson correlations between ring- width index and monthly climate variables. Panels (**b-d**) give statistics for time windows tested in *climwin*, where window open and close indicate months prior to current August, and cells across the lower diaganol indicate single-month tests (akin to panel **a**). Panels (**b**) and (**c**) give values of linear and quadratic terms for each time window, and (**d**) gives the $\Delta AIC$ for each. The time window with the minimum $\Delta AIC$ (0-3 months prior to August, or May-July; black circles), was identified as the optimal window. Panel (**e**) shows the correlation of individual-level residuals to PET, with the function fit in *climwin*. Finally, panel (**f**) shows GLS model output, where $PET$ was a candidate driver variable (along with $PPT$; $DBH$ not included in this model). Plotted are responses of species for which $PET$ was included in the top model, with best-fit polynomials plotted with solid lines when both first- and second-order terms are signficant, DASHED lines when only one term is signficant, and DOTTED lines when neither is signficant. Transparent ribbons indicate 95% confidence intervals.](tables_figures/quilt_comparison.png)

@@

**Interactive effects of $DBH$ and climate were found for #% of species-variable combinations**. 
Interactions were significant for over half of $DBH$-precipitation variable interactions for all three growth metrics (*51% for all*; Table S5).
The majority (#%) of these interactions were positive, indicating that larger trees generally respond more positively (or less negatively) to precipitation or it's frequency (Fig. **6?**).
Interactions between temperature variables and $DBH$ were significant for *~40%* of cases considered (Table S5). 
The majority (#%) of these interactions were positive, indicating that larger trees generally respond less negatively (or more positively) to $T_{max}$ or $PET$ (Fig. **6?**).

### 2020-08-27 doc/manuscript/MEE_manuscript_SI.Rmd
update SI fig numbers

>>>>

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
allometries_table <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climate_interactions_summary.csv", stringsAsFactors = FALSE, check.names = FALSE)
#kable(allometries_table, format = "latex", booktabs = TRUE,   escape = F)%>%
#  kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```
